---
title: "Reproducible Research - Project 1"
author: "Chieh-Ju Lin"
date: "09/01/2017"
output: 
  html_document: 
    keep_md: yes
---

Load the data and check it

```{r}
data <- read.csv("activity.csv")
head(data)
tail(data)
str(data)
```



* Let's first check some simple plot and statistics

Format the data and calculate the total number of steps taken per day

```{r}
library(dplyr)
data_1 <- na.omit(data)
by_date <- group_by(data_1, date)
total_steps <- summarise(by_date, sum(steps, na.rm = TRUE))
names(total_steps) <- c("date", "steps")
total_steps
```

Build histogram

```{r}
library(ggplot2)
g <- ggplot(total_steps, aes(steps))
g <- g + geom_histogram(breaks = seq(0, 25000, by = 2000), color = "black")
g <- g + labs(title = "Histogram of Total Number of Steps per Day", x = "Steps")
g
```

The mean of the total number of steps taken per day is:

```{r}
mean(total_steps$steps)
```

The median of the total number of steps taken per day is:

```{r}
median(total_steps$steps)
```



* Now, let's check the average daily activity pattern

Format the data

```{r}
data_2 <- group_by(data, interval)
by_interval <- summarise(data_2, average_steps = mean(steps, na.rm = TRUE))
by_interval
```

Build a time series plot of the 5-minute interval and the average number of steps taken

```{r}
g <- ggplot(by_interval, aes(interval, average_steps))
g <- g + geom_line(aes(x = interval, y = average_steps), size = 1)
g <- g + labs(x = "Interval", y = "Average steps", title = "Average Daily Activity Pattern")
g
```

Looks like on average people have highest number of steps during the morning

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
by_interval[which.max(by_interval$average_steps), ]
```



* Let's handle the missing values

How many missing values in the dayaset?

```{r}
summary(data)
sum(is.na(data$steps))
```

Let's simply use the mean for each 5-minute interval to replace the missing values here and create a new dataset

```{r}
# Create a new column with average steps by interval
rep <- rep(by_interval$average_steps, 61)
data_new <- mutate(data, average_steps = rep)

# Replace NA with average steps using for loop
for (i in 1:17568){
        if (is.na(data_new$steps[i]) == TRUE){
                data_new$steps[i] <- data_new$average_steps[i]
        }
}

# Check NAs
sum(is.na(data_new$steps))

# Drop the new column
data_new <- data_new[, c(1:3)]
head(data_new)
```

Let's make a new histogram for the new dataset

```{r}
# Group steps by date
data_3 <- group_by(data_new, date)
by_date_new <- summarise(data_3, steps = sum(steps))
head(by_date_new)

# Build histogram
g <- ggplot(by_date_new, aes(steps))
g <- g + geom_histogram(breaks = seq(0, 25000, by = 2000), color = "black")
g <- g + labs(title = "Histogram of Total Number of Steps per Day", x = "Steps")
g
```

The mean of the total number of steps taken per day for the new dataset is:

```{r}
mean(by_date_new$steps)
```

The median of the total number of steps taken per day for the new dataset is:

```{r}
median(by_date_new$steps)
```

As we can see, the mean steps maintain the same, while the median steps increase a little bit  
  
  
  
* Let's further check if there are differences in activity patterns between weekdays and weekends

We first create a new factor variable in the dataset with two levels ¡V ¡§weekday¡¨ and ¡§weekend¡¨ indicating whether a given date is a weekday or weekend day

```{r}
head(data_new)
Sys.setlocale("LC_TIME", "English")
data_new$day_of_week <- weekdays(as.Date(data_new$date))
for (i in 1:17568) {
        if (data_new$day_of_week[i] %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")){
                data_new$day_of_week[i] <- "Weekday"
        }
        else{ 
                data_new$day_of_week[i] <- "Weekend"
        }
}
table(data_new$day_of_week)
```

Then let's format the dataset and make a panel plot

```{r}
# Group the data by day_of_week and interval
data_4 <- group_by(data_new, day_of_week, interval)
by_interval_2 <- summarise(data_4, steps = mean(steps))
head(by_interval_2)

# Make a panel plot containing a time series plot
g <- ggplot(by_interval_2, aes(x = interval, y = steps))
g <- g + geom_line(size = 1) + facet_grid(day_of_week~.)
g <- g + labs(x = "Interval", y = "Average steps", title = "Average Daily Activity Pattern")
g
```

During the weekdays, people tend to have the highest number of steps in the morning  
And during the weekends, the activity pattern have multiple peaks
